Group 2 Final Project Report

### 2013-12-19 reportLatex/paper.tex
pending conclusion updated

\title{Group 2 Final Project Report (revB1218)}
\author{Jay Guild, Nathan Miller, Matthew L Weber}

@@

\section{Introduction}
\IEEEPARstart{A}{cceleration} of a cryptographic algorithms on an embedded device remains an outstanding challenge. New System on Chip (SoC) FPGA technologies provide more options to offload or accelerate the cryptographic computations, however, integrating these technologies into existing designs can be challenging. Similarly, migrating designs between different SoC FPGA platforms can also require a significant investment of time and resources. Even though the system functionality and underlying algorithms remain unchanged, the design teams are forced to re-invent their designs for each new platform.

@@

\section{Related Work}
Cryptography is the process of encrypting information using an algorithm and either a public key (asymmetric encryption) or a private key (symmetric encryption) that makes the information unreadable unless the data is passed through a decryption algorithm that utilizes a private key that only the receiver holds\cite{HIPPA}. Several keys and algorithms have been developed under the IEEE standard for public key cryptography \cite{IEEESTD} and studied in several different environments for effectiveness and speed. It has been found that over networks with cryptography implemented that the limiting factor for most systems in speed is the network itself so using software to handle cryptography has not helped to speed up processing time much \cite{GUTMANN}. With speeds always increasing on communication buses this will change soon though. The largest bonuses in the recent past have been seen when adding a co-processor or FPGA to handle cryptography. By adding a co-processor or FPGA effectiveness is increased because computations can be done in parallel and logic can be handled in fewer steps\cite{GUERON}. Computations are also hard wired and coded into circuits and that creates a security that software in a single dedicated processor can’t match because that software can be manipulated by outside code\cite{GUTMANN}. A SoC FPGA that can process cryptography very fast using parallelization in a low cost implementation can resolve any issues that using a single processor software solution creates.

\section{Why is this interesting...}
In the last two years various industry leading programmable logic companies have invested millions in creating consolidated system on a chip (SoC) products. Fusing mainstream processor technology (primarily ARM-based cores) with the latest fabric technology, the result has been the creation of highly configurable embedded devices. This is very relevant to bleeding edge signal processing developments (cryptography/video/audio/RF), where a general purpose processor is commonly paired with a programmable device for initial prototyping and in some cases the final product. The latest SoC FPGA technology allows a consolidation that has largely eliminated many interface design activities and has encouraged more of a accelerator/coprocessor style of development. The key to this flexibility is having the SoC execute code that may tap into a fabric logic to parallelize an operation or having SoC completely hand off data to the fabric (leaving the SoC free to perform other operations).

Parallel operations in an FPGA using incorruptible LUTs and unidentifiable return data also has the positive effect of added security for sensitive systems such as the one being implemented here.  When a possible intrusion to the system would reach the encoding of the FPGA there is a blockage in communication that would not allow the intruder to see what operations are being done to the incoming data to the microprocessor from the FPGA.  To be able to garner data an intruder would need to be able to monitor data prior to processing and that is only done on the outgoing side from the originator.  There is also no way for an intruder to manipulate the FPGA to try and detect the encoding due to it’s nature of having permanent operations that are a physical architecture inside the chip.  This makes this type of implementation very useful in security sensitive communication systems.

@@

\section{Project Overview}
Our original project proposal suggested using a framework called Open Crypto Framework (OCF)\cite{OCF2013}.  It turned out OCF was deprecated and not functional with newer Linux kernels.  Instead the Cryptodev development now provides a comparable capability along with a set of test applications used for validation and benchmarking.  This works out to fulfill the Crypto API and TestApp portions of the design.  The SHA1 entity is based on an Opencore core
.
As originally proposed, the architecture Figure \ref{fig_archBlkH} consists of a software kernel driver API, user space application, as well as a programmable logic (PL) entity.  Using these components, an initial implementation yielded a functional implementation of the SHA1 algorithm using the PL to compute the hash results.

@@

\section{Design Description}
The design consists of the components outlined in Figure \ref{fig_archBlkDetailed}.  Each of these is expanded in more detail in the sections that follow below.
\subsection{Hardware}
The Xilinx Zynq SoC FPGA 702 development kit was used as our target hardware. We acquired one development kit for our project and coordinated integration efforts to share the workload of only having one kit.  ISE tool support was included in the university's 14.6 license and the ISE EDK builder was used to pull together the embedded design. The Zynq ARM core Linux maturity was adequate in the Xilinx 14.6 release and Xilinx had some pretty good examples of interfacing between the fabric and processor. \cite{XIL2013} \cite{XILGNU2013}   The Zynq consists of two parts, the processing system 7 (PS7) and the programmable logic (PL). The PL might also be referred to as the “fabric” in the sections that follow.  The Zynq development kit was interfaced to via Ethernet and serial. With a base ARM load-set executing a Linux Kernel configured for modular development and capable of loading the PL images.
\subsection{Software}
For the software operating environment a full embedded Linux 3.12 Kernel and filesystem were used, allowing use of existing userspace applications and the Linux kernel’s existing infrastructure for cryptography.\cite{WIKPE2013} \cite{OCF2013} This configuration allowed the project to maximize reuse of the baseline algorithms already present in open source software and focus on gathering benchmark comparison data of our hardware accelerated approach.  The Xilinx Zynq board ARM OS load-set was based on Xilinx GIT tree source [14], with the Kernel/Rootfs being compiled into loadable binaries using the Buildroot \cite{BUILDROOT2013} embedded build system. We also added the Linux Cryptodev 1.6 support to the kernel and utilized the Cryptodev sha speed test app as part of our Buildroot rootfs.  The first stage bootloader (FSBL) was created using the Xilinx 14.6 EDK toolflow to export a XSDK FSBL project for our custom PL design.
\subsubsection{TestApp}
The test application used for executing the benchmarking of the baseline software algorithms on the SoC was based on the sha1\_speed application provided by the cryptodev-linux package.  It performed a number of block size test scenarios and captured throughput performance of each.

This application was also modified during integration (sha\_test) to allow generating of arbitrarily sized msg sizes to validate hash function behavior and buffer handling code.  It was also configured to send known msg sequences to validate against known hash results.  This was especially important when validating the behaviors were as expected between the single message and the multimessage scenarios (in multi-message the hash from the previous message is the seed to the next hash computation).
\subsubsection{Cryptodev Framework (Kernel Module)}
The Cryptodev framework’s linux device driver is designed to provide userspace applications access to Linux kernel crypto algorithms.  In our scenarios it’s used as is to provide initial test metrics against existing kernel software based algorithms.  As well as later it was used with the Zynq PL version of the SHA1 Opencore algorithm to run comparable test sequences to generate performance metrics.
\subsubsection{CryptoAPI (Kernel Interface)}
The kernel standard interface to all kernel crypto algorithms.  This was not modified as part of our project and instead a previous algorithm implemented was used as a template to help us conform to the existing Linux kernel Crypto API standard.
\subsubsection{PL SHA1 Interface Module}
The interface to the PL algorithm was implemented through a memory mapped set of registers (via a AXI interconnect to the PL).  The registers provided 32bit word access to control and status information.  They also provided the resulting hash and a method for data input.  The software interface code implemented a behavior that serviced the algorithm user logic entity based on register status feedback.  Currently the PL side of this interface does not use interrupts and requires the software to poll a register for status.

This software interface received its data from the crypto API framework (Userspace --$>$ kernel) in blocks and then translated that data into what was needed to feed the SHA1 algorithm in the PL.  The diagram Appendix Figure C-1 shows the interactions between software and the PL.  The userspace TestApp would send data to the Cryptodev driver that would intern call the Kernel Crypto API that would call the sha1\_update function for each block of data.  When the last block is passed, the kernel API calls the sha1\_final.  In sha1\_final, the tail is appended before hashing the final bits of data.  Followed by the retrieval of the hash before the function returns.  
\subsection{Programmable Logic}
The Programmable Logic (PL) portion of our project consisted of three sub-components: (1) an AXI slave, (2) the SHA1 logic, and (3) a user-logic entity that implemented the register interface and interfaced with the SHA1 logic (user\_logic.vhd).  In addition to these three synthesized components, several simulation-only entities were created to verify the correctness of our design and to extract preliminary performance metrics.
Development of the PL entities were performed using the Xilinx 14.6 ISE/XPS tools.
Each of the main sub-components are explored in the following sub-sections.
\subsubsection{Glue (PCORE)}
As shown in Figure \ref{fig_archBlkDetailed}, an AXI bus connects the ARM core to the PL. The Xilinx CORE Generator System \cite{COREGEN} was used to create an AXI slave called 'glue' (glue.vhd).  We did not alter the generated 'glue' component for this project.
\begin{figure}[ht]
\centering
\includegraphics[width=2.5in]{Sha1_Core_Block_Diagram.png}
\caption{SHA-1 Core}
\label{fig_algoCore}
\end{figure} 
\subsubsection{SHA1 Core}
Our SHA1 core shown in Figure \ref{fig_algoCore} was based upon an open-source SHA1 core written by Arif Nugroho\cite{SHAOPEN}. After studying its implementation, we created a test-bench for the core (sha\_tb/sha1\_tb.vhd and sha\_tb/sha1\_tb\_driver.vhd) to both verify that it produced the correct hash results and to confirm our assumptions about its usage.

Starting from a working core saved us a great deal of development time. This particular core also had the distinct advantage of having no external dependencies beyond the standard IEEE libraries. However, the standalone nature of the core also posed several challenges. The core imposed strict timing requirements for both supplying message data and extracting the hash results. Loading sixteen words of message data into the core required a new word be supplied to the msg input signal every clock cycle for sixteen consecutive clock cycles. There was no way to delay message input. Similarly extracting the five word hash result required reading the hash signal for five consecutive clock cycles. The core user had exactly one clock cycle to begin capturing the hash results once the core's 'valid' signal became active.  Failure to do observe these requirements would produce incorrect hash results.

@@

These challenges of SHA1 core were addressed in two ways. First, a FSM was added to the glue logic to insure the timing requirements for supplying message values and retrieving hash results were satisfied (see the next section). Secondly, the SHA1 core was modified to save a hash result only when new message data had previously been supplied.  This small modification to the core effectively stalled its progress until new data had been supplied by the user.
\subsubsection{User-logic SW Interface (PL)}
As discussed in the "PL SHA1 Interface Module" section, the PL SHA1 Interface software running on the ARM core expects to read and write to memory-mapped registers.  To achieve this capability, the Xilinx CORE Generator System was used to create the initial user\_logic component that defines those registers, and connects them to the AXI slave.  This user\_logic component was altered for this project to fit our requirements.  Specifically, it was modified to define behaviors for the different registers, and to define an FSM for interfacing with the SHA1 core.

@@

One possible improvement would be to enhance the user\_logic entity to use interrupts to signal when the software should transition states.  This would reduce the overhead of bus transactions required to maintain status and free the ARM processor up to execute other tasks.  It would also slightly increase throughput as a few additional clock cycles might become available.
\begin{figure}[ht]
\centering
\includegraphics[width=2.5in]{UserLogicStateDiagram.png}
\caption{SHA-1 State Machine}
\label{fig_algoFSM}
\end{figure}
As discussed in the previous section, the slightly modified SHA1 core had specific timing requirements associated with supplying the message block and retrieving the hash results.  To accommodate these requirements a FSM called ‘user\_fsm’ was created as shown in Figure \ref{fig_algoFSM}.   This FSM idles in the ‘algoWait’ state until either a 1 value (hash new message block) or a 2 value (hash next message block) is written into the (unfortunately named) ‘rst’ register.  After proceeding through the ‘algoBegin’ state (discussed later), it then supplies 16 words of data into the SHA1 core from the corresponding user\_logic registers (dataXfer).  It then waits until the hash operation completes which is signaled when the SHA1 core raises the ‘valid’ signal (hashWait).  Once the hash is ready, it retrieves the five words of hash data from the core and stores them into the appropriate user\_logic registers (saveHash), before transitioning back to the ‘algoWait’ state.

One of the user\_logic design goals was to gracefully handle stale entries in the rst register.  Early version of user\_fsm would immediately begin a new round of hashing unless the software cleared the rst register soon after setting it.  Imposing this requirement on software was unacceptable.  At the same time, we wanted user\_logic to handle software that, for whatever reason, wrote the same value to the rst register over any number of cycles.  In both situations we wanted the user\_fsm to only begin a new round of hashing when software wrote a “fresh” 1 or 2 value to the rst register.

To accommodate both design goals, a new\_op signal was defined, and the user\_fsm was modified to only leave the algoWait state if new\_op was set.  The process that updated the rst register in response to AXI activity sets new\_op to 1 only when the rst register was NOT updated in the in the previous clock cycle or a different value was written to the rst register in the previous clock cycle.  This prior state history was maintained via two new signals updated each clock cycle: p\_reg\_write\_sel and p\_Bus2IP\_Data.  The new\_op signal was cleared when the user\_fsm transitioned to the ‘algoBegin’ state.

The overall operation of SHA1 PL design is shown in a flow diagram in the Appendix Figure C-1. All flow starts from the (top) userspace application running on the ARM.  The boxes in green are the kernel function calls that implement the custom SHA1 interfacing to the PL (sha1\_update/final as discussed in Section: “PL SHA1 Interface Module”).  The boxes in yellow are interactions with the PL register map (FSM).  The boxes in red are return/final states.  Bringing this all together, the diagram shows the software control access and PL status interactions, as well as the data load and hash retrieval steps.

\section{Experimental/Testing Results}
\subsection{Simulation}
\begin{figure}[ht]
\centering
\includegraphics[width=2.5in]{Test_Benches.png}
\caption{Test bench coverage}
\label{fig_TB}
\end{figure} 
Our simulation effort for this project focused on two components: the open-source SHA1 core and the modified user\_logic component.  As outlined above, we simulated the SHA1 core to insure that it produced the correct hash results, to validate our understanding of the core’s operation and to confirm that stalling the core worked properly.  Simulation of the user\_logic component validated the proper functioning of the user\_fsm, and was also used to estimate overall performance and extract the relative amount of time spent in the primary user\_logic states (the yellow boxes in Appendix Figure C-1).

For each simulation effort a distinct test-bench was created as shown in Figure \ref{fig_TB}.  To make the user-logic test-bench more readily extend-able the logic was divided into a single generic “test-harness” component (user\_logic\_tb.vhd) , and a collection of individual test components.  The test-harness executed each test separately and performed a reset of the user\_logic logic prior to running the next test.  Tests consisted of validating the hash result of a single message block, multiple independent message blocks, and chained message blocks.  The Linux command-line tool ‘sha1sum’\cite{SHASUM}  was used to calculate the expected hash result for each of the test messages, and these hash results were further validated using online SHA1 calculators\cite{SHASUMONLINE}.

@@

The following subsections describe each of the user\_logic tests and present the performance metrics extracted from each.  Result data is detailed in Appendix A: Table A-1.
\subsubsection{Test1-Bus access}
\begin{figure}[ht]
\centering
\includegraphics[width=2.5in]{Test1.jpg}
\caption{Test1 Runtime}
\label{fig_test1Runtime}
\end{figure} 
The initial simulation was used as a benchmark to give a metric as to the actual time it would take to write to and read from the FPGA after sufficient data had been received at the microprocessor.  Four bytes of data were written to the output registers, sent to the FPGA, and read back without encoding them.  The bytes were then compared with their initial values to verify there was no corruption.  An initial metric of 17 ns for overall process time, 8 ns for data transfer, 8 ns for verification, and 1 ns idle time was established to determine overhead per 4 bytes of data received as shown in Figure \ref{fig_test1Runtime}.  This step validated that the signals, registers, and encryption algorithm were working properly with the IP2Bus and provided a baseline for all other metrics.
\subsubsection{Test2 - Single block}
\begin{figure}[ht]
\centering
\includegraphics[width=2.5in]{Test2.jpg}
\caption{Test2 Runtime}
\label{fig_test2Runtime}
\end{figure} 
The second test in the simulation was to send 44 bytes of data through the SHA1 hash as a single block and validate the generated hash.  This test was set to validate an entire block of data passed through the SHA-1 core.  The write time increased disproportionately to the amount of data sent due to wait time in between writing to the registers and the passing of data to the SHA-1 core.  Hash time was added in for the time in hash process and the validation time increased proportionally to the amount of data sent.  See Appendix A: Table A-1 for detailed metrics used to create Figure \ref{fig_test2Runtime}.
\subsubsection{Test3 - Single block twice}
\begin{figure}[ht]
\centering
\includegraphics[width=2.5in]{Test3.jpg}
\caption{Test3 Runtime}
\label{fig_test3Runtime}
\end{figure}
The third test in the simulation was to send 44 bytes of data through SHA1 hash twice, each time computing a single block and validating the generated hash.  This test was set to validate the proper function of writing to the registers multiple times with a proper hash calculated with each block sent.  The write time and validation time doubled as expected while the hash time was optimized and showed a better performance than Test 2.  See Appendix A: Table A-1 for detailed metrics used to create Figure \ref{fig_test3Runtime}. 
\subsubsection{Test4 - Multi block}
\begin{figure}[ht]
\centering
\includegraphics[width=2.5in]{Test4.jpg}
\caption{Test4 Runtime}
\label{fig_test4Runtime}
\end{figure} 
The fourth test in the simulation was to send 88 bytes of data through the SHA1 hash and have the algorithm process it as two blocks (multi-message).  This test was to validate the proper function of running multiple blocks through the SHA-1 core while maintaining an appended hash.  Write time stayed the same as expected while validation time was halved due to only one hash result needing to be verified, unlike Test 3.  Hash time only increased slightly and was due to the need to carry over the hash from the first block instead of the hard coded initial hash values.  See Appendix A: Table A-1 for detailed metrics used to create Figure \ref{fig_test4Runtime}.

@@

Another reality of the SHA1 algorithm was that it is inherently hard to speed up through parallelization (it being a sequential block hash).  You really instead end up counting instructions and optimizing the approach to supplying the data.  Compared to a non-block cipher that could easily be parallelized and accelerated using the method this project used. 
\subsection{Target Hardware}
1) For our first scenario we wanted to verify that our kernel, rootfs, and cryptodev kernel module were functional.  We cross built a kernel and rootfs with cryptodev support (kernel module and test apps), and then loaded them onto the sdcard and booted the Zynq board.

Next, we executed the sha\_speed test and recorded the results.  The standard test sequence included loading the cryptodev kernel module to give us access to the Kernel Crypto APIs and the respective sha1\_asm or sha1\_zynq algorithm module.  
The kernel in this case was configured to use the default sha1\_arm hash algorithm, which is an ASM optimized implementation for ARM.  The results of this scenario are part of the graph under performance metrics (figure X).

>>>>

\title{Group 2 Final Project Report}
\author{Jay Guild, Nathan Miller, Matthew L Weber}

@@

\section{Introduction}
\IEEEPARstart{A}{cceleration} of a cryptographic algorithm on an embedded device remains an outstanding challenge. New System on Chip (SoC) FPGA technologies provide more options to offload or accelerate the cryptographic computations, however, integrating these technologies into existing designs can be challenging. Similarly, migrating designs between different SoC FPGA platforms can also require a significant investment of time and resources. Even though the system functionality and underlying algorithms remain unchanged, the design teams are forced to re-invent their designs for each new platform.

@@

\section{Related Work}
Cryptography is the process of encrypting information using an algorithm and either a public key (asymmetric encryption) or a private key (symmetric encryption) that makes the information unreadable unless the data is passed through a decryption algorithm that utilizes a private key that only the receiver holds\cite{HIPPA}. Several keys and algorithms have been developed under the IEEE standard for public key cryptography \cite{IEEESTD} and studied in several different environments for effectiveness and speed. It has been found when using cryptography over network packets, the limiting factor for most systems is the network itself and not the cryptography software implementation being used\cite{GUTMANN}. With speeds always increasing on communication buses this will change soon though. The largest bonuses in the recent past have been seen when adding a co-processor or FPGA to handle cryptography. By adding a co-processor or FPGA the effectiveness is increased because computations can be done in parallel and logic can be handled in fewer steps\cite{GUERON}. Computations are also hard coded into circuits and are less amendable to software based attacks\cite{GUTMANN}.

\section{Why is this interesting...}
In the last two years various industry leading programmable logic companies have invested millions in creating consolidated system on a chip (SoC) products. Fusing mainstream processor technology (primarily ARM-based cores) with the latest fabric technology, the result has been the creation of highly configurable embedded devices. This is very relevant to bleeding edge signal processing developments (cryptography/video/audio/RF), where a general purpose processor is commonly paired with a programmable device for initial prototyping and in some cases the final product. The latest SoC FPGA technology allows a consolidation that has largely eliminated many interface design activities and has encouraged more of a accelerator/coprocessor style of development. The key to this flexibility is having the SoC execute code that may tap into the fabric logic to parallelize an operation or having SoC completely hand off data to the fabric (leaving the SoC free to perform other operations).  There is also no way for an intruder to manipulate the FPGA to try and detect the encoding due to it’s nature of having permanent operations that are a physical architecture inside the chip.  This makes this type of implementation very useful in security sensitive communication systems.

@@

\section{Project Overview}
Our original project proposal suggested using a framework called Open Crypto Framework (OCF)\cite{OCF2013}.  It turned out OCF was deprecated and not functional in newer Linux kernels.  Instead the Cryptodev development now provides a comparable capability along with a set of test applications used for validation and benchmarking.  This package provides compliant Crypto API and TestApp portions that we levered in our project design.  Lastly the SHA1 programmable logic entity was based on an Opencore core.

As was originally proposed, the architecture in Figure \ref{fig_archBlkH} consists of a software kernel driver API, user space application, as well as a programmable logic (PL) entity.  Using these components, an initial implementation yielded a functional implementation of the SHA1 algorithm using the PL to compute the hash results.

@@

\section{Design Description}
The design consists of the components outlined in Figure \ref{fig_archBlkDetailed}.  Each of these is expanded in more detail in the sections that follow below.
\subsection{Hardware}
The Xilinx Zynq SoC FPGA 702 development kit was used as our target hardware. We acquired one development kit for our project and coordinated integration efforts to share the workload of only having one kit.  ISE tool support was included in the university's 14.6 license and the ISE EDK builder was used to pull together the embedded design. The Zynq ARM core Linux maturity was adequate in the Xilinx 14.6 release and Xilinx had mature examples of interfacing between the fabric and processor. \cite{XIL2013} \cite{XILGNU2013}   The Zynq consists of two parts: the processing system 7 (PS7) and the programmable logic (PL). The PL is also referred to as the "fabric" in the sections that follow.  Communication to the Zynq development kit was provided via Ethernet and serial interfaces.  The ARM operating system load-set executes a Linux Kernel configured for modular development and is capable of loading the PL images.
\subsection{Software}
For the software operating environment a full embedded Linux 3.12 Kernel and filesystem were used, allowing use of existing userspace applications and the Linux kernel’s existing infrastructure for cryptography.\cite{WIKPE2013} \cite{OCF2013} This configuration allowed the project to maximize reuse of the baseline algorithms already present in open source software and focus on gathering benchmark comparison data of our hardware accelerated approach.  The Xilinx Zynq board ARM OS load-set was based on Xilinx GIT tree source [14], with the Kernel/Rootfs being compiled into loadable binaries using the Buildroot \cite{BUILDROOT2013} embedded build system. We also added the Linux Cryptodev 1.6 support to the kernel and utilized the Cryptodev sha speed test app as part of our Buildroot rootfs.  The first stage bootloader (FSBL) was created using the Xilinx 14.6 EDK toolflow to export a XSDK FSBL project for our custom PL design.
\subsubsection{TestApp}
The test application used for executing the benchmarking of the baseline software algorithms on the SoC was based on the sha1\_speed application provided by the cryptodev-linux package.  It performed a number of block size test scenarios and captured throughput performance of each.

We modified this application to allow generating of arbitrarily sized msg sizes.  Which allowed us to validate custom hash function behavior and buffer handling code.  It was also configured to send known msg sequences to validate against known hash results.  This was important when validating that the behaviors were as expected in the single message and the multi-message scenarios (in multi-message the hash from the previous message is the seed to the next hash computation).
\subsubsection{Cryptodev Framework (Kernel Module)}
The Cryptodev framework's Linux device driver provides Userspace applications access to Linux kernel crypto algorithms.  In our scenarios it is used as is to provide initial test metrics against existing kernel software based algorithms.  It was also used against a Zynq PL version of the SHA1 Opencore algorithm to run comparable test sequences and generate performance metrics.
\subsubsection{CryptoAPI (Kernel Interface)}
The Crypto API is kernel's standard interface to all kernel cryptographic module functions.  The API was not modified as part of our project.  We designed our cryptographic algorith to interface in with the existing Linux kernel Crypto API standard.
\subsubsection{PL SHA1 Interface Module}
This software interface received its data from the crypto API framework (Userspace --$>$ kernel) in blocks and then translated that data into what was needed to feed the SHA1 algorithm in the PL.  The diagram Appendix Figure C-1 shows the interactions between software and the PL.  The Userspace TestApp would send data to the Cryptodev driver that would in-turn call the Kernel Crypto API to invoke the sha1\_update function for each block of data.  When the last block is supplied, the kernel API calls the sha1\_final function.  In sha1\_final, a message tail is appended before hashing the final bits of data.  Followed by the retrieval of the hash before the function returns.

The interface to the PL algorithm was implemented through a memory mapped set of registers (via a AXI interconnect to the PL).  The registers provided 32bit word access to control the PL, retrieve status information, provided the resulting hash and input the message data.  The software interface used these registers to interface to the PL.  Currently the PL side of this interface does not use interrupts and requires the software to poll a register for status.

\subsection{Programmable Logic}
The Programmable Logic (PL) portion of our project consisted of three sub-components: (1) an AXI slave, (2) the SHA1 logic, and (3) a user-logic entity that implemented the register interface and interfaced with the SHA1 logic (user\_logic.vhd).  In addition to these three synthesized components, several simulation-only entities were created to verify the correctness of our design and to extract preliminary performance metrics.
The Xilinx 14.6 ISE/XPS tools were used to develop the PL entities.  Each of these sub-components are discussed in the following sub-sections.
\subsubsection{Glue (PCORE)}
As shown in Figure \ref{fig_archBlkDetailed}, an AXI bus connects the ARM core to the PL. The Xilinx CORE Generator System \cite{COREGEN} was used to create an AXI slave called 'glue' (glue.vhd).  We did not alter the generated 'glue' component for this project.
\begin{figure}[ht]
\centering
\includegraphics[width=2.5in]{Sha1_Core_Block_Diagram.png}
\caption{SHA-1 Core}
\label{fig_algoCore}
\end{figure} 
\subsubsection{SHA1 Core}
Our SHA1 core shown in Figure \ref{fig_algoCore} was based upon an open-source SHA1 core written by Arif Nugroho\cite{SHAOPEN}. After studying its implementation, we created a test-bench for the core (sha\_tb/sha1\_tb.vhd and sha\_tb/sha1\_tb\_driver.vhd) to both verify that it produced the correct hash results and to confirm our assumptions about its usage.

Starting from a working core saved us a great deal of development time. This particular core also had the distinct advantage of having no external dependencies beyond the standard IEEE libraries. However, the standalone nature of the core also posed several challenges. The core imposed strict timing requirements for both supplying message data and extracting the hash results. Loading sixteen words of message data into the core required a new word be supplied to the msg input signal every clock cycle for sixteen consecutive clock cycles. There was no way to delay message input. Similarly extracting the five word hash result required reading the hash signal for five consecutive clock cycles. The core's user had exactly one clock cycle to begin capturing the hash results once the core's 'valid' signal became active.  Failure to do observe these requirements would cause incorrect hash results.

@@

These challenges were addressed in two ways. First, a Finite State Machine (FSM) was added to the glue logic to insure the timing requirements of the SHA1 core.   Secondly, the SHA1 core was modified to save a hash result only when new message data had previously been supplied.  This small modification to the core effectively stalled its progress until new data had been supplied by the user.
\subsubsection{User-logic SW Interface (PL)}
As discussed in the "PL SHA1 Interface Module" section, the PL SHA1 Interface software running on the ARM core expects to read and write to memory-mapped registers.  To achieve this capability the Xilinx CORE Generator System was used to create the initial user\_logic component which defines those registers, and connects them to the AXI slave.  This user\_logic component was altered for this project to fit our requirements.  Specifically, it was modified to define behaviors for the different registers, and to define an FSM for interfacing with the SHA1 core.

@@

One possible improvement would be to enhance the user\_logic entity to use interrupts to signal when the software should transition states.  This would reduce the overhead of bus transactions required to maintain status and free the ARM processor up to execute other tasks.  It would also slightly increase throughput as a few additional clock cycles might become available.
\begin{figure}[ht]
\centering
\includegraphics[width=2.5in]{UserLogicStateDiagram.png}
\caption{SHA-1 State Machine}
\label{fig_algoFSM}
\end{figure}

As discussed in the previous section, the slightly modified SHA1 core had specific timing requirements associated with supplying the message block and retrieving the hash results.  To accommodate these requirements a FSM called 'user\_fsm' was created [Figure \ref{fig_algoFSM}].   This FSM idles in the 'algoWait' state until either a 1 value (hash new message block) or a 2 value (hash next message block) is written into the (unfortunately named) 'rst' register.  After proceeding through the 'algoBegin' state (discussed later), it then supplies sixteen words of data into the SHA1 core from the corresponding user\_logic registers (dataXfer).  It then waits until the hash operation completes which is signaled when the SHA1 core raises the 'valid' signal (hashWait).  Once the hash is ready, it retrieves the five words of hash data from the core and stores them into the appropriate user\_logic registers (saveHash), before transitioning back to the 'algoWait' state.

One of the user\_logic design goals was to gracefully handle stale entries in the rst register.  Early versions of user\_fsm would immediately begin a new round of hashing unless the software cleared the 'rst' register soon after setting it.  Imposing this requirement on software was unacceptable.  At the same time, we wanted user\_logic to handle software that, for whatever reason, wrote the same value to the 'rst' register for any number of consecutive cycles.  In both situations we wanted the user\_fsm to only begin a new round of hashing when software wrote a "fresh" 1 or 2 value to the 'rst' register.

To accommodate both design goals, a new\_op signal was defined, and the user\_fsm was modified to only leave the algoWait state if new\_op was set.  The process that updated the 'rst' register in response to AXI activity sets new\_op to 1 only when the 'rst' register was NOT updated in the previous clock cycle, or when a different value was written to the 'rst' register in the previous clock cycle.  This history of the 'rst' register and it's prior reset state was maintained in two new signals: p\_reg\_write\_sel and p\_Bus2IP\_Data.  The new\_op signal was cleared when the user\_fsm transitioned to the 'algoBegin' state.

The overall operation of SHA1 PL design is shown in a flow diagram in the Appendix Figure C-1. All flow starts from the (top) userspace application running on the ARM.  The boxes in green are the kernel function calls that implement the custom SHA1 interfacing to the PL (sha1\_update/final as discussed in Section: "PL SHA1 Interface Module").  The boxes in yellow are interactions with the PL register map (FSM).  The boxes in red are return/final states.  All together, the diagram shows the software control access and PL status interactions, as well as the data load and hash retrieval steps.

\section{Experimental/Testing Results}
\subsection{Simulation}
\begin{figure}[ht]
\centering
\includegraphics[width=2.5in]{Test_Benches.png}
\caption{Test bench coverage}
\label{fig_TB}
\end{figure} 
Our simulation effort for this project focused on two components: the open-source SHA1 core and the modified user\_logic component.  As outlined above, we simulated the SHA1 core to insure that it produced the correct hash results.  These results validated our understanding of the core's operation and confirmed that our modifications to stall the core worked properly.  Simulation of the user\_logic component validated the proper functioning of the user\_fsm, and was also used to estimate overall performance and extract the relative amount of time spent in the primary user\_logic states (the yellow boxes in Appendix Figure C-1).

Distinct test-benches were created for each simulation effort [Figure \ref{fig_TB}].  To make the user-logic test-bench more readily extend-able the logic was divided into a single generic "test-harness" component (user\_logic\_tb.vhd), and a collection of individual test components.  The test-harness executed each test separately and performed a reset of the user\_logic logic prior to running the next test.  The Linux command-line tool 'sha1sum'\cite{SHASUM}  was used to calculate the expected hash result for each of the test messages, and these hash results were further validated using online SHA1 calculators\cite{SHASUMONLINE}.

@@

The following subsections describe each of the user\_logic tests and present the performance metrics extracted from each.  Result data is detailed in Appendix A: Table A-1.
\subsubsection{Test1-Bus access}
\begin{figure}[ht]
\centering
\includegraphics[width=2.5in]{Test1.jpg}
\caption{Test1 Runtime}
\label{fig_test1Runtime}
\end{figure}
The initial simulation was used as a benchmark to give a metric as to the actual time it would take to write to and read from the FPGA after sufficient data had been received at the microprocessor.  Four bytes of data were written to the output registers, sent to the FPGA, and read back without encoding them.  The bytes were then compared with their initial values to verify there was no corruption.  An initial metric of 17ns for overall process time, 8ns for data transfer, 8ns for verification, and 1ns idle time was established to determine overhead per 4bytes of data received as shown in Figure \ref{fig_test1Runtime}.  This step validated that the signals and registers were working properly with the IP2Bus and provided a baseline for all other metrics.
\subsubsection{Test2 - Single block}
\begin{figure}[ht]
\centering
\includegraphics[width=2.5in]{Test2.jpg}
\caption{Test2 Runtime}
\label{fig_test2Runtime}
\end{figure} 
The second test in the simulation was to send 44bytes of data through the SHA1 hash as a single block and validate the generated hash.  This test was set to validate an entire block of data passed through the SHA-1 core.  The write time increased disproportionately to the amount of data sent due to wait time in between writing to the registers and the passing of data to the SHA-1 core.  Hash time was added in for the time in hash process and the validation time increased proportionally to the amount of data sent.  See Appendix A: Table A-1 for detailed metrics used to create Figure \ref{fig_test2Runtime}.
\subsubsection{Test3 - Single block twice}
\begin{figure}[ht]
\centering
\includegraphics[width=2.5in]{Test3.jpg}
\caption{Test3 Runtime}
\label{fig_test3Runtime}
\end{figure}
The third test in the simulation was to send 44bytes of data through SHA1 hash twice, each time computing a single block and validating the generated hash.  This test was set to validate the proper function of writing to the registers multiple times with a proper hash calculated with each block sent.  The write time and validation time doubled as expected while the hash time was optimized and showed a better performance than Test 2.  See Appendix A: Table A-1 for detailed metrics used to create Figure \ref{fig_test3Runtime}. 
\subsubsection{Test4 - Multi block}
\begin{figure}[ht]
\centering
\includegraphics[width=2.5in]{Test4.jpg}
\caption{Test4 Runtime}
\label{fig_test4Runtime}
\end{figure} 
The fourth test in the simulation was to send 88bytes of data through the SHA1 hash and have the algorithm process it as two blocks (multi-message).  This test was to validate the proper function of running multiple blocks through the SHA-1 core while maintaining an appended hash.  Write time stayed the same as expected while validation time was halved due to only one hash result needing to be verified, unlike Test 3.  Hash time only increased slightly and was due to the need to carry over the hash from the first block instead of the hard coded initial hash values.  See Appendix A: Table A-1 for detailed metrics used to create Figure \ref{fig_test4Runtime}.

@@

Another reality of the SHA1 algorithm was that it is inherently hard to speed up through parallelization (it being a sequential block hash).  You really instead end up counting instructions and optimizing the approach to supplying the data.  Compared to a non-block cipher that could easily be parallelized and accelerated using the method this project used. 
\subsection{Target Hardware}
Development on the Zynq development kit proceeded as a series of scenarios.
1) For our first scenario we wanted to verify that our kernel, rootfs, and cryptodev kernel module were functional.  We cross built a kernel and rootfs with cryptodev support (kernel module and test apps), and then loaded them onto the sdcard and booted the Zynq board.

Next, we executed the sha\_speed test and recorded the results.  The standard test sequence included loading the cryptodev kernel module to give us access to the Kernel Crypto APIs and the respective sha1\_asm or sha1\_zynq algorithm module.  
The kernel in this case was configured to use the default sha1\_arm hash algorithm, which is an ASM optimized implementation for ARM.  The results of this scenario are part of the graph under performance metrics [Figure \ref{fig_algoPerf}]).
